---
title: "R Project Submission HEK"
author: "Hamza Elkoutri"
date: "4/30/2022"
output:
  html_document: default
  pdf_document: default
runtime: shiny
---

This document ,developped by Hamza ELKOUTRI, details a "Trading calculator" application made on R Shiny. The strategy consists of several technical indicators : EMA, SMA, SMI and lastly a comparison with the benchmark. We'll go by this code block by block and explain it through comments.

## I-Necesseray packages

Firstly you'll need to install all the packages used in this code and import them. Excpetionally for the package "quantmod", due to bugs in Yahoo Finance, please install it from this source : "remotes::install_github("joshuaulrich/quantmod",force=TRUE)" and "remotes::install_github("joshuaulrich/quantmod@358-getsymbols-new.session",force=TRUE)"


```{r packages, collapse=TRUE,message=FALSE}
library(xts)
library(shiny)
library(shinythemes)
library(data.table)
library(TTR)
library(quantmod)
library(PerformanceAnalytics)
```

## II-Functions used for this project

We'll define all the functions block by block.

### Extracting the stock data from yfinance 


```{r extractions, message=FALSE, collapse=TRUE}
Get_DB<-function(Ticker,DateFrom,DateTo){
  DBT<-getSymbols(Ticker, src = 'yahoo', from = DateFrom, to = DateTo,auto.assign = FALSE)
  DBT$Simple_Ret<- 100*diff(Cl(DBT))/lag(Cl(DBT), 1)
  DBT$Log_Ret<- 100*diff(log(Cl(DBT)))
  DBT$Simple_Ret[is.na(DBT$Simple_Ret)] <- 0
  DBT$Log_Ret[is.na(DBT$Log_Ret)] <- 0
  colnames(DBT)<-c("DB.Open","DB.High","DB.Low","DB.Close","DB.Volume","DB.Adjusted","DB.Simple_Ret","DB.Log_Ret")
  return(DBT)
}
```

### Building technical indicators

Now we will use built in packages for trading indicators

#### Exponential Moving Average Indicator(EMA)

This calculates fast diff(20-50 EMA) that gives in and out market

```{r EMA_Calculations, message=FALSE, collapse=TRUE}
Trigger_EMA<-function(DB,period){
  ema_period_db <- EMA(Cl(DB), n = period)
  return(ema_period_db)
}

Crossover_EMA<-function(DB){
  ema_period_A_db <- EMA(Cl(DB), n = 20)
  ema_period_B_db <- EMA(Cl(DB), n = 50)
  ema_db_trigger <- Lag(
    ifelse(Lag(ema_period_A_db) < Lag(ema_period_B_db) & ema_period_A_db > ema_period_B_db,1,
           ifelse(Lag(ema_period_A_db) > Lag(ema_period_B_db) & ema_period_A_db < ema_period_B_db,-1,0)))
  ema_db_trigger[is.na(ema_db_trigger)] <- 0
  chartSeries(DB,TA=c(addVo(),addEMA(20,col="red"),addEMA(50,col = "pink"),addTDI(),addVolatility()))
   legend('left', col = c('green','red','pink'),
          legend = c(paste('Ticker'),'EMA_20','EMA_50'), lty = 1, bty = 'n',
          text.col = 'white', cex = 0.8)
  return(ema_db_trigger)
}
```

#### Simple Moving Average Indicator(SMA) and Stochastic Momentum Index(SMI)

We do the same with SMA and SMI

```{r SMA_SMI_Calculations, message=FALSE, collapse=TRUE}
Trigger_SMA<-function(DB,period){
  sma_period_db <- SMA(Cl(DB), n = period)
  return(sma_period_db)
}

Crossover_SMA<-function(DB){
  sma_period_A_db <- Trigger_SMA(DB,20)
  sma_period_B_db <- Trigger_SMA(DB,50)
  sma_db_trigger <- Lag(
    ifelse(Lag(sma_period_A_db) < Lag(sma_period_B_db) & sma_period_A_db > sma_period_B_db,1,
           ifelse(Lag(sma_period_A_db) > Lag(sma_period_B_db) & sma_period_A_db < sma_period_B_db,-1,0)))
  sma_db_trigger[is.na(sma_db_trigger)] <- 0
  chartSeries(DB,TA=c(addVo(),addSMA(20,col="red"),addSMA(50,col = "pink"),addTDI(),addVolatility()))
  legend('left', col = c('green','red','pink'),
         legend = c(paste('Ticker'),'SMA_20','SMA_50'), lty = 1, bty = 'n',
         text.col = 'white', cex = 0.8)
  
  return(sma_db_trigger)
}

Trigger_SMI<-function(DB){
  smi_db <- SMI(HLC(DB),
                n = 13, nFast = 2, nSlow = 25, nSig = 9)
  return(smi_db)
}

Crossover_SMI<-function(DB){
  smi_db <- Trigger_SMI(DB)
  smi_db_trigger <- Lag(
    ifelse(Lag(smi_db[,1]) < Lag(smi_db[,2]) & smi_db[,1] > smi_db[,2],1, 
           ifelse(Lag(smi_db[,1]) > Lag(smi_db[,2]) & smi_db[,1] < smi_db[,2],-1,0)))
  smi_db_trigger[is.na(smi_db_trigger)] <- 0
  chartSeries(DB,TA=c(addVo(),addSMI(),addTDI(),addVolatility()))
  legend('left', col = c('green'),
         legend = c(paste('Ticker')), lty = 1, bty = 'n',
         text.col = 'white', cex = 0.8)
  return(smi_db_trigger)
}
```

### Building trading strategies

We'll create trading stretagies using the triggers we just built

```{r Trading_Strategies, message=FALSE, collapse=TRUE}
Strategy_EMA<-function(DB){
  ema_A_db <- Trigger_EMA(DB,20)
  ema_B_db <- Trigger_EMA(DB,50)
  ema_db_trigger <- Crossover_EMA(DB)
  ema_db_strat <- ifelse(ema_db_trigger > 1,0,1)
  for (i in 1 : length(Cl(DB))) {
    ema_db_strat[i] <- ifelse(ema_db_trigger[i] == 1,1,ifelse(ema_db_trigger[i] == -1,0,ema_db_strat[i-1]))
  }
  ema_db_strat[is.na(ema_db_strat)] <- 1
  ema_db_strat_final <- cbind(ema_A_db, ema_B_db, ema_db_trigger, ema_db_strat)
  colnames(ema_db_strat_final) <- c('EMA(20)','EMA(50)','EMA SIGNAL','EMA POSITION')
  return(ema_db_strat_final)
}

Strategy_SMA<-function(DB){
  sma_A_db <- Trigger_SMA(DB,20)
  sma_B_db <- Trigger_SMA(DB,50)
  sma_db_trigger <- Crossover_SMA(DB)
  sma_db_strat <- ifelse(sma_db_trigger > 1,0,1)
  for (i in 1 : length(Cl(DB))) {
    sma_db_strat[i] <- ifelse(sma_db_trigger[i] == 1,1,ifelse(sma_db_trigger[i] == -1,0,sma_db_strat[i-1]))
  }
  sma_db_strat[is.na(sma_db_strat)] <- 1
  sma_db_strat_final <- cbind(sma_A_db, sma_B_db, sma_db_trigger, sma_db_strat)
  colnames(sma_db_strat_final) <- c('SMA(20)','SMA(50)','SMA SIGNAL','SMA POSITION')
  return(sma_db_strat_final)
}

Strategy_SMI<-function(DB){
  smi_db <- Trigger_SMI(DB)
  smi_db_trigger <- Crossover_SMI(DB)
  smi_db_strat <- ifelse(smi_db_trigger > 1,0,1)
  for (i in 1 : length(Cl(DB))) {
    smi_db_strat[i] <- ifelse(smi_db_trigger[i] == 1,1,ifelse(smi_db_trigger[i] == -1,0,smi_db_strat[i-1]))
  }
  smi_db_strat[is.na(smi_db_strat)] <- 1
  smi_db_strat_final <- cbind(smi_db[,1],smi_db[,2],smi_db_trigger,smi_db_strat)
  colnames(smi_db_strat_final) <- c('SMI','SMI(S)','SMI SIGNAL','SMI POSITION')
  return(smi_db_strat_final)
}
```

This additional function is only to be used on the shiny app later. It basically facilates the reading of the data, instead of having binary signals

```{r Trading_Strategies_UI, message=FALSE, collapse=TRUE}
Strategy_All_UI<-function(DB){
  ema_A_db <- Trigger_EMA(DB,20)
  ema_B_db <- Trigger_EMA(DB,50)
  ema_db_trigger <- Crossover_EMA(DB)
  
  ema_db_strat_UI <- ifelse(ema_db_trigger > 1,"Not holding","Holding")
  for (i in 1 : length(Cl(DB))) {
    ema_db_strat_UI[i] <- ifelse(ema_db_trigger[i] == 1,"Holding",ifelse(ema_db_trigger[i] == -1,"Not holding",ema_db_strat_UI[i-1]))
  }
  ema_db_strat_UI[is.na(ema_db_strat_UI)] <- "Holding"
  
  ema_db_trigger_UI <- Lag(
    ifelse(Lag(ema_A_db) < Lag(ema_B_db) & ema_A_db > ema_B_db,"Buy",
           ifelse(Lag(ema_A_db) > Lag(ema_B_db) & ema_A_db < ema_B_db,"Sell","Neutral")))
  ema_db_trigger_UI[is.na(ema_db_trigger_UI)] <- "Neutral"
  
  sma_A_db <- Trigger_SMA(DB,20)
  sma_B_db <- Trigger_SMA(DB,50)
  sma_db_trigger <- Crossover_SMA(DB)
  
  sma_db_strat_UI <- ifelse(sma_db_trigger > 1,"Not holding","Holding")
  for (j in 1 : length(Cl(DB))) {
    sma_db_strat_UI[j] <- ifelse(sma_db_trigger[j] == 1,"Holding",ifelse(sma_db_trigger[j] == -1,"Not holding",sma_db_strat_UI[j-1]))
  }
  sma_db_strat_UI[is.na(sma_db_strat_UI)] <- "Holding"
  
  sma_db_trigger_UI <- Lag(
    ifelse(Lag(sma_A_db) < Lag(sma_B_db) & sma_A_db > sma_B_db,"Buy",
           ifelse(Lag(sma_A_db) > Lag(sma_B_db) & sma_A_db < sma_B_db,"Sell","Neutral")))
  sma_db_trigger_UI[is.na(sma_db_trigger_UI)] <- "Neutral"
  
  db_strat_final <- cbind(data.frame(Dates=index(DB),coredata(DB)),ema_db_trigger_UI, ema_db_strat_UI,sma_db_trigger_UI,sma_db_strat_UI)
  colnames(db_strat_final) <- c("Dates","DB.Open","DB.High","DB.Low","DB.Close","DB.Volume","DB.Adjusted","DB.Simple_Ret","DB.Log_Ret",'EMA TRIGGER','EMA POSITION','SMA TRIGGER','SMA POSITION')
  return(db_strat_final)
}
```

### Results and performance

We now build all the results that are based on all of the functions that we defined previously

```{r Results_Strategies, message=FALSE, collapse=TRUE}
Ticker_Return<-function(DB){
  ret_db <- diff(log(Cl(DB)))
  benchmark_db <- ret_db
  return(benchmark_db)
}

Results_EMA<-function(DB){
  benchmark_db<-Ticker_Return(DB)
  ema_db_strat_final<-Strategy_EMA(DB)
  ema_db_trigger<-Crossover_EMA(DB)
  ema_db_strat<-ema_db_strat_final[,4]
  ema_db_ret <- benchmark_db*ema_db_strat
  ema_db_ret_commission_adj <- ifelse((ema_db_trigger == 1|ema_db_trigger == -1) & ema_db_strat != Lag(ema_db_trigger), (benchmark_db-0.05)*ema_db_strat, benchmark_db*ema_db_strat)
  ema_db_comp <- cbind(ema_db_ret, ema_db_ret_commission_adj, benchmark_db)
  colnames(ema_db_comp) <- c('EMA','EMA Commission Adj','Ticker Benchmark')
  #charts.PerformanceSummary(ema_db_comp, main =paste('EMA Performance'))
  return(ema_db_comp)
}

Results_SMA<-function(DB){
  benchmark_db<-Ticker_Return(DB)
  sma_db_strat_final<-Strategy_SMA(DB)
  sma_db_trigger<-Crossover_SMA(DB)
  sma_db_strat<-sma_db_strat_final[,4]
  sma_db_ret <- benchmark_db*sma_db_strat
  sma_db_ret_commission_adj <- ifelse((sma_db_trigger == 1|sma_db_trigger == -1) & sma_db_strat != Lag(sma_db_trigger), (benchmark_db-0.05)*sma_db_strat, benchmark_db*sma_db_strat)
  sma_db_comp <- cbind(sma_db_ret, sma_db_ret_commission_adj, benchmark_db)
  colnames(sma_db_comp) <- c('SMA','SMA Commission Adj','Ticker Benchmark')
  #charts.PerformanceSummary(sma_db_comp, main =paste('SMA Performance'))
  return(sma_db_comp)
}

Results_SMI<-function(DB){
  benchmark_db<-Ticker_Return(DB)
  smi_db_strat_final<-Strategy_SMI(DB)
  smi_db_trigger<-Crossover_SMI(DB)
  smi_db_strat<-smi_db_strat_final[,4]
  smi_db_ret <- benchmark_db*smi_db_strat
  smi_db_ret_commission_adj <- ifelse((smi_db_trigger == 1|smi_db_trigger == -1) & smi_db_strat != Lag(smi_db_trigger), (benchmark_db-0.05)*smi_db_strat, benchmark_db*smi_db_strat)
  smi_db_comp <- cbind(smi_db_ret, smi_db_ret_commission_adj, benchmark_db)
  colnames(smi_db_comp) <- c('SMI','SMI Commission Adj','Ticker Benchmark')
  #charts.PerformanceSummary(smi_db_comp, main =paste('SMI Performance'))
  return(smi_db_comp)
}
```

Lastly, these functions are used to calculate the profit made for an intial investment on this stock

```{r Profits, message=FALSE, collapse=TRUE}
Profit_EMA<-function(DB,invest){
  i<-0
  Buy<-invest
  Growth <-0
  ema_db_strat_final<-Strategy_EMA(DB)
  while(i<=length((ema_db_strat_final$`EMA SIGNAL`))){
    if(isTRUE((ema_db_strat_final$`EMA POSITION`)[i]==1))
    {if(isTRUE((ema_db_strat_final$`EMA SIGNAL`)[i]==1)){
      
      Buy<-Buy-as.numeric(Cl(DB)[i])
    }
      Growth<-Growth+as.numeric(DB$DB.Simple_Ret[i])
    }
    if(isTRUE((ema_db_strat_final$`EMA SIGNAL`)[i]==-1)){
      Buy<-Buy+as.numeric(Cl(DB)[i])}
    i<-i+1}
  return(paste('With ',invest,'$(USD) investment, profit of this Strategy EMA on this stock: ',invest*Growth/100,'$(USD)'))
}

Profit_SMA<-function(DB,invest){
  i<-0
  Buy<-invest
  Growth <-0
  sma_db_strat_final<-Strategy_SMA(DB)
  while(i<=length((sma_db_strat_final$`SMA SIGNAL`))){
    if(isTRUE((sma_db_strat_final$`SMA POSITION`)[i]==1))
    {if(isTRUE((sma_db_strat_final$`SMA SIGNAL`)[i]==1)){
      Buy<-Buy-as.numeric(Cl(DB)[i])
    }
      Growth<-Growth+as.numeric(DB$DB.Simple_Ret[i])
    }
    if(isTRUE((sma_db_strat_final$`SMA SIGNAL`)[i]==-1)){
      Buy<-Buy+as.numeric(Cl(DB)[i])}
    i<-i+1}
  return(paste('With ',invest,'$(USD) investment, profit of this Strategy SMA on this stock: ',invest*Growth/100,'$(USD)'))
}

Profit_SMI<-function(DB,invest){
  i<-0
  Buy<-invest
  Growth <-0
  smi_db_strat_final<-Strategy_SMI(DB)
  while(i<=length((smi_db_strat_final$`SMI SIGNAL`))){
    if(isTRUE((smi_db_strat_final$`SMI POSITION`)[i]==1))
    {if(isTRUE((smi_db_strat_final$`SMI SIGNAL`)[i]==1)){
      Buy<-Buy-as.numeric(Cl(DB)[i])
    }
      Growth<-Growth+as.numeric(DB$DB.Simple_Ret[i])
    }
    if(isTRUE((smi_db_strat_final$`SMI SIGNAL`)[i]==-1)){
      Buy<-Buy+as.numeric(Cl(DB)[i])}
    i<-i+1}
  return(paste('With ',invest,'$(USD) investment, profit of this Strategy SMI on this stock: ',invest*Growth/100,'$(USD)'))
}
```

## III-Application on SHINY

This code defines the UI and all the calls of the functions we defined, in order to have a user friendly interface that create our trading strategies.

```{r Shiny_APP, message=FALSE, collapse=TRUE}
ui <- fluidPage(
  tags$img(src = "https://emlyon.brightspace.com/d2l/lp/navbars/6606/theme/viewimage/495707/view?v=20.22.4.36719",height = 50, width = 50),
  headerPanel("Trading calculator (by HEK)"),
  theme = shinytheme("yeti"),
  sidebarLayout(
  sidebarPanel(
     h4("Choose your parameters:"),
     textInput(inputId = "ticker", label="Enter ticker symbol:",placeholder = "AAPL, TSLA, TWTR...",value="TSLA"),
     dateInput(inputId = "Start_date",label="Choose starting date:",max=Sys.Date(),value=Sys.Date()-365),
     dateInput(inputId = "End_date",label="Choose starting date:",max=Sys.Date(),value=Sys.Date()),
     numericInput(inputId = "Capital",label="Enter your inital investement in USD($):",min=0,value=100)),
      mainPanel(
        tabsetPanel(
          tabPanel("Overall View",dataTableOutput(outputId = 'raw_extract')),
          tabPanel("EMA 20-50 Analysis", h4("Graphical representation:"),plotOutput(output="Plot_EMA"),h4("Returns annualized:"),tableOutput(outputId = 'Result_EMA')),
          tabPanel("SMA 20-50 Analysis", h4("Graphical representation:"),plotOutput(output="Plot_SMA"),h4("Returns annualized:"),tableOutput(outputId = 'Result_SMA')),
          tabPanel("SMI Analysis", h4("Graphical representation:"),plotOutput(output="Plot_SMI"),h4("Returns annualized:"),tableOutput(outputId = 'Result_SMI')),
          tabPanel("Performance Analysis",plotOutput(output="Perf_EMA"),plotOutput(output="Perf_SMA"),plotOutput(output="Perf_SMI")),
          tabPanel("Profit forecast",h4("Profit following EMA:"),textOutput(outputId="Profit_EMA"),h4("Profit following SMA:"),textOutput(outputId="Profit_SMA"),h4("Profit following SMI:"),textOutput(outputId="Profit_SMI"))
          )
      )
  )
)

server <- function(input, output) {
  output$raw_extract <-renderDataTable(Strategy_All_UI(Get_DB(input$ticker,input$Start_date,input$End_date)))
  output$Result_EMA <- renderTable(table.AnnualizedReturns(Results_EMA(Get_DB(input$ticker,input$Start_date,input$End_date))))
  output$Plot_EMA<-renderPlot(Crossover_EMA(Get_DB(input$ticker,input$Start_date,input$End_date)))
  output$Result_SMA <- renderTable(table.AnnualizedReturns(Results_SMA(Get_DB(input$ticker,input$Start_date,input$End_date))))
  output$Plot_SMA<-renderPlot(Crossover_SMA(Get_DB(input$ticker,input$Start_date,input$End_date)))
  output$Result_SMI <- renderTable(table.AnnualizedReturns(Results_SMI(Get_DB(input$ticker,input$Start_date,input$End_date))))
  output$Plot_SMI<-renderPlot(Crossover_SMI(Get_DB(input$ticker,input$Start_date,input$End_date)))
  output$Perf_EMA<-renderPlot(charts.PerformanceSummary(Results_EMA(Get_DB(input$ticker,input$Start_date,input$End_date)), main =paste('EMA Performance')))
  output$Perf_SMA<-renderPlot(charts.PerformanceSummary(Results_SMA(Get_DB(input$ticker,input$Start_date,input$End_date)), main =paste('SMA Performance')))
  output$Perf_SMI<-renderPlot(charts.PerformanceSummary(Results_SMI(Get_DB(input$ticker,input$Start_date,input$End_date)), main =paste('SMI Performance')))
  output$Profit_EMA<-renderText(Profit_EMA(Get_DB(input$ticker,input$Start_date,input$End_date),input$Capital))
  output$Profit_SMA<-renderText(Profit_SMA(Get_DB(input$ticker,input$Start_date,input$End_date),input$Capital))
  output$Profit_SMI<-renderText(Profit_SMI(Get_DB(input$ticker,input$Start_date,input$End_date),input$Capital))
  
  }

shinyApp(ui = ui, server = server)
```

## General Remarks

This concludes our code. This code has been tested, in case there's a problem compiling it, please reach to me at : hamza.elkoutri@edu.em-lyon.com

Since this is a shiny app, another ".r" file will be included in the package to make compilation easier.